The growth of children's semantic and phonological networks: insight from 10 languages

### 2019-09-20 journal/ms.Rmd
revisions

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky; @swingley2018].

@@

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. Thus, according to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. According to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al (2009).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@@

Second, we study vocabularies measured using both comprehension and production. Previous studies have found differences between these vocabularies in terms of their content and rate of acquisition [@benedict1979; @fenson94; @bates1995]. These differences may reflect the fact that comprehension and production do not share the same constraints. For instance, whereas comprehension depends on the ease with which words are stored and accessed, production depends, additionally, on the ease with which words are articulated, e.g., shorter words are produced earlier [@braginsky]. By investigating comprehension-based vocabularies, we assess the extent to which the network growth mechanism captures general learning patterns beyond the specific constraints of production.

@@

## Age of acquisition
For each word in the CDI data, we compute the proportion of children who understand or produce the word at each month. Then we fit a logistic curve to these proportions and determined when the curve crosses 0.5, i.e., the age at which at least 50% of children know the word. We take this point in time to be each word’s age of acquisition [@goodman2008; @braginsky].
<!--We obtained these nouns in ten languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the hand-checked translation equivalents provided by @braginsky2016, allowing us to use the English association norms across languages. Table \@ref(tab:stats) gives an overview of the data used. -->

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for the dataset we used.",
      col.names = c("Language",  "Nouns", "Ages", "N", "Nouns", "Ages","N")) %>%
  add_header_above(c("", "Comprehension" = 3, "Production" = 3)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

@@

plot_correlation_prod_1 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), 
                                  aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

plot_correlation_prod_2 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

@@

plot_correlation_comp_1 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

plot_correlation_comp_2 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age")

@@

```{r}
##Test
growth_preds <- feather::read_feather("../saved_data/growth_preds.feather")

@@

plot_growth_preds <- ggplot(subset(growth_preds, Test=='Individual'), aes(x = model , y = mean)) +
  geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model, linetype=measure),
  #geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.3)))+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7) + #,+
        #axis.text.x = element_text(size=7.5, angle=45)) +
  scale_y_continuous(breaks =c(0,0.2))  +
  #scale_y_discrete(breaks =NULL)
  ylab("Estimate") +xlab("Mechanism")

@@

```{r growthPred, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Evaluation of growth scenarios (EXT: externally-driven, INT: internally-driven) for both semantic and phonological networks. Each point represents the mean of the posterior distribution of the growth parameter, with ranges representing 95\\% credible intervals. Positive values mean that learning proceeds according to the predictions of the growth scenario, whereas negative values mean that learning proceeds in opposition to the predictions of the growth scenario.", fig.width=8, fig.height=3}

@@

plot_growth_preds

@@

```{r}
##Test
static_preds <- feather::read_feather("../saved_data/static_preds.feather")

@@

plot_static_preds <- ggplot(subset(static_preds, Test=='Combined'), aes(x = predictor , y = estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., col = predictor, linetype=measure),
  #geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.3)))+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7) + #,+
        #axis.text.x = element_text(size=7.5, angle=45)) +
  scale_y_continuous(breaks =c(-1,0,1))  +
  #scale_y_discrete(breaks =NULL)
  ylab("Estimate") +xlab("Mechanism")

@@

```{r staticPred, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the regression model in each language. Results are shown for both production and comprehension data. Ranges indicate 95\\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).", fig.width=8, fig.height=3}

@@

plot_static_preds

@@

plot_static_preds_all <- ggplot(static_preds_all, aes(x = predictor, y = Estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., y = Estimate, col = predictor, linetype=Test),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  facet_wrap(~measure)  +
  coord_flip() +
  guides(colour=FALSE)+
  scale_colour_solarized() +
  theme_few()+
  theme(aspect.ratio = 0.7)

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008 for a similar research strategy].

@@

The findings for the new predictors were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition in both comprehension and production data and across languages, endorsing results for nouns across a variety of analyses [@goodman2008; @braginsky; @roy2015]. Word length is more predictive for production than comprehension (and this difference is very clear in the global model), replicating previous work [@braginsky]. Thus, word length seems to reflect the effects of production's constraints rather than comprehension's constraints, i.e., longer words are harder to articulate but they may not be significantly more difficult to store and access.

@@

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking low level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994] and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that show how learning general phonotactics patterns create "well-worn paths" which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017].

@@

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT [@beckage]. To illustrate, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

@@

>>>>

What factors shape vocabulary learning over the course of early childhood? To investigate this question, scientists have adopted multiple research strategies, from conducting controlled laboratory experiments [e.g. @markman90] to analyzing dense corpora capturing language learning in context [e.g., @roy2015]. One prominent strategy consists in documenting the timeline of words' acquisition, and studying the properties that make words easy or hard to learn [e.g., @goodman2008; @huttenlocher1991]. For example, @goodman2008 found that, within a lexical category (e.g., nouns), higher parental frequency is associated with earlier learning. Researchers have studied the role of several other factors such as word length and the mean length of utterances in which the word occurs [e.g., @braginsky2019; @swingley2018].

@@

```{r growth, fig.cap = "Illustration of the two growth scenarios. Filled grey circles represent known words (Internal) at a certain point in time. The empty, numbered circles represent words that have not yet been learned (External) and which are candidates to enter the lexicon next. The identity of the word that is going to be learned depends on the growth scenario.  Here the squares indicate the node that drives growth in each scenario and the hand pointer indicates which word is likely to be learned. For INT, the utility of a candidate, external node is the average degree (i.e., number of links) of the internal nodes that it would attach to. In this simplified example, candidate node 1 would connect to an internal node with 3 connections, thus we have $u_{INT}(node_1)= 3$. As for candidate node 2, it would connect to internal nodes that have only one connection each, making an average of 1, i.e., $u_{INT}(node_2)= 1$. According to INT, node 1 is more likely to enter the lexicon. For EXT, the utility of a candidate node is its degree in the entire network. In our example, candidate node 1 has 2 connections in total whereas candidate node 2 has 5 connections. So we have $u_{EXT}(node_1)= 2$ and $u_{EXT}(node_2)= 5$. Thus, according to EXT, node 2 is more likely to enter the lexicon next. This figure is based on an example from Hills et al. (2009).", fig.align = "center", out.width = "300px"}
knitr::include_graphics("figs/growth3.png")
```

@@

Second, we study vocabularies measured using both comprehension and production. Previous studies have found differences between these vocabularies in terms of their content and rate of acquisition [@benedict1979; @fenson94; @bates1995]. These differences may reflect the fact that comprehension and production do not share the same constraints. For instance, whereas comprehension depends on the ease with which words are stored and accessed, production depends, additionally, on the ease with which words are articulated, e.g., shorter words are produced earlier [@braginsky2019]. By investigating comprehension-based vocabularies, we assess the extent to which the network growth mechanism captures general learning patterns beyond the specific constraints of production.

@@

## Age of acquisition
For each word in the CDI data, we compute the proportion of children who understand or produce the word at each month. Then we fit a logistic curve to these proportions and determined when the curve crosses 0.5, i.e., the age at which at least 50% of children know the word. We take this point in time to be each word’s age of acquisition [@goodman2008; @braginsky2019].
<!--We obtained these nouns in ten languages: Croatian, Danish, English, French, Italian, Norwegian, Russian, Spanish, Swedish, and Turkish. We used the subset of nouns that had entries in the Florida Association Norms (see below). Since these norms are available only in English, we used the hand-checked translation equivalents provided by @braginsky2016, allowing us to use the English association norms across languages. Table \@ref(tab:stats) gives an overview of the data used. -->

@@

kable(statistics, format = "latex", escape = FALSE, booktabs = TRUE,
      linesep = "", format.args = list(big.mark = ","),
      caption = "Statistics for the dataset we used. The ages are in months.",
      col.names = c("Language",  "Nouns", "Ages", "N", "Nouns", "Ages","N")) %>%
  add_header_above(c("", "Comprehension" = 3, "Production" = 3)) %>%
  column_spec(1, bold = TRUE) %>%
  kable_styling(position = "center")

@@

cor_phon_len_edit2 <- data_all %>%
  group_by(language, measure) %>%
  summarise(phon_len = cor(phono_deg, length)) %>%
  group_by(measure) %>%
  summarise(ave = mean(phon_len))

@@

plot_correlation_prod_1 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), 
                                  aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

plot_correlation_prod_2 <- ggplot(subset(data_long, measure == "produces" & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="produces" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

@@

plot_correlation_comp_1 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

plot_correlation_comp_2 <- ggplot(subset(data_long, measure == "understands" & aoa < 35 & data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ predictor)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, measure=="understands" & correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=25, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

@@

```{r}
##Test

#Colorblind-friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

@@

plot_growth_preds <- ggplot(subset(growth_preds, Test=='Individual'), aes(x = model , y = mean)) +
  geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model, linetype=measure),
  #geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.5)))+
  scale_colour_manual(values=cbPalette)+
  theme_few()+
  theme(aspect.ratio = 0.7,
        axis.text = element_text(size=12)) +
  scale_y_continuous(breaks =c(0,0.2))  +
  #scale_y_discrete(breaks =NULL)
  ylab("Estimate") +xlab("Mechanism")

@@

```{r growthPred, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Evaluation of growth scenarios (EXT: externally-driven, INT: internally-driven) for both semantic and phonological networks. Each point represents the mean of the posterior distribution of the growth parameter, with ranges representing 95\\% credible intervals. Positive values mean that learning proceeds according to the predictions of the growth scenario, whereas negative values mean that learning proceeds in opposition to the predictions of the growth scenario.", fig.width=10, fig.height=4}

@@

plot_growth_preds + theme(legend.position="bottom")

@@

```{r}
##Test
static_preds <- feather::read_feather("../saved_data/coefs_all.feather")

@@

plot_static_preds <- ggplot(subset(static_preds, Test=='Combined'), aes(x = predictor , y = estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., col = predictor, linetype=measure),
  #geom_pointrange(aes(ymin = hdi_lower, ymax = hdi_upper, col = model),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  coord_flip() +
  facet_wrap( ~ language, ncol=5) +
  guides(colour=FALSE, linetype = guide_legend(override.aes = list(size=0.6)))+
  scale_colour_manual(values=cbPalette)+
  theme_few()+
  theme(aspect.ratio = 0.7,
        axis.text = element_text(size=12)) +
  scale_y_continuous(breaks =c(-1,0,1))  +
  #scale_y_discrete(breaks =NULL)
  ylab("Estimate") +xlab("Mechanism")

@@

```{r staticPred, out.width = "\\textwidth", fig.pos = "!h", fig.cap = "Estimates of the relative contribution of each predictor of AoA in the regression model in each language. Results are shown for both production and comprehension data. Ranges indicate 95\\% confidence intervals. Positive values indicate a positive relationship (e.g. longer words tend to have a higher AoA), while negative values indicate a negative relationship (e.g. words with higher frequency tend to have a lower AoA).", fig.width=10, fig.height=4}

@@

plot_static_preds + theme(legend.position="bottom")

@@

plot_static_preds_all <- ggplot(static_preds_all, aes(x = predictor, y = Estimate)) +
  geom_pointrange(aes(ymin = X2.5.., ymax = X97.5.., y = Estimate, col = predictor, linetype=Test),
                  position = position_dodge(width = .5),
                  size = 0.5,
                  fatten = 0.5)+
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed")+
  facet_wrap(~measure)  +
  coord_flip() +
  guides(colour=FALSE)+
  scale_colour_manual(values=cbPalette)+
  theme_few()+
  theme(aspect.ratio = 0.7)

@@

Above we showed that the way semantic and phonological information is structured in the learning environment contributes to noun learning (via EXT) across languages. However, we know that other factors influence learning as well [e.g., @braginsky2019]. Next we investigated how semantic and phonological connectivity interact with two other factors. The first one is word frequency, a well studied factor shown to predict the age of acquisition in a reliable fashion [e.g. @goodman2008]. The second factor is word length, which was shown to correlate with phonological connectivity: Shorter words are more likely to have higher connectivity [@pisoni1985; @vitevitch2005].

@@

For word length, we counted the number of phonemes in our generated IPA transcription. For word frequency, we used the frequency estimates from @braginsky2019 where unigram counts were derived based on CHILDES corpora in each language [@macwhinney2014]. Although these frequency counts use transcripts from independent sets of children, they are based on large samples, and this allows us to average out possible differences between children and the specificities of their input [see @goodman2008 for a similar research strategy].

@@

The findings for the new predictors were as follows. Overall, frequency is the largest and most consistent predictor of age of acquisition in both comprehension and production data and across languages, endorsing results for nouns across a variety of analyses [@goodman2008; @braginsky2019; @roy2015]. Word length is more predictive for production than comprehension (and this difference is very clear in the global model), replicating previous work [@braginsky2019]. Thus, word length seems to reflect the effects of production's constraints rather than comprehension's constraints, i.e., longer words are harder to articulate but they may not be significantly more difficult to store and access.

@@

In the phonological case, network growth according to EXT is also compatible with a scenario whereby children are tracking low level statistical patterns, e.g., high probability sound sequences. Indeed, connectivity in the phonological network is inherently correlated with phonotactic probability [@vitevitch1999]. That is, highly connected words tend to be made of frequent sound sequences. Children are sensitive to local phonotactic regularities [@jusczyk1994] and this sensitivity might lead them to learn higher-probability words more easily [@storkel2001]. This explanation is supported by computational simulations that show how learning general phonotactics patterns create "well-worn paths" which allow the models to represent several distinct but phonologically neighboring words [@dell1993; @takac2017; @siew2013]. More generally, there is a growing interest in investigating precisely how the local patterns acquired through statistical learning may give rise to the global network orgnization [For a review, see @karuza2016].

@@

This study investigated the class of nouns in isolation --- following previous studies investigating the early semantic and phonological network [@hills2009; @storkel2009]. We could ask if studying one class separtaly is a legitimate research strategy. In other words, would word classes (such as nouns, verbs and function words) be acquired relatively differently, or would they interact substantially to the extent that it becomes unreasonable to study each class separately?

We know that different word classes follow different time trajectories. For instance, in the early stages of development, nouns tend to be acquired at a higher pace than predicates and function words [@bates1994]. Research has shown that this difference cannot be trivially attributed to differences in the degree to which these classes are present in the input; if anything, verbs and function words are often more frequent in the input than nouns [e.g., @gentner1982].  @goodman2008 found an effect of frequency on the age of acquisition within --- not across --- classes. Further, a recent work by @braginsky2019 tested a large number of predictors, besides frequency, and found that these predictors do not influence the acquisition of word classes in the same way. For example, the acquisition of nouns was found to be most influenced by frequency and concreteness, whereas the acquisition of function words was most influenced by word length.

These patterns support the hypothesis that different word classes have different pathways of learning, thus making it worthwhile to study each class separately. This hypothesis, however, does not preclude all sorts of inter-class interaction. For example, @dautriche2015 found that the acquisition of similar-sounding words ("bih"-"dih") is easier if these words belong to different syntactic categories (e.g., a noun and a verb).

This work shares a number of limitations with previous studies using similar research strategy and datasets. Chief among these limitations is the fact that the age of word acquisition is computed using different children at different ages (due to the fact that available CDI data is mainly cross-sectional). Such a measure has been shown to be valid and reliable [@fenson94], and has allowed researchers to study important aspects of word learning [@hills2009; @stella2017; @storkel2009; @goodman2008; @braginsky2019]. In our case, the use of cost-effective cross-sectional data has allowed us to leverage large-scale studies across several languages. That said, it is important to remember that this type of data can only inform us about the learning trajectory of the "average" child. Although our study endorses, overall, the externally-driven account of network growth, this does not mean individual children never use some variant of INT or some combination of both INT and EXT [@beckage]. To illustrate, some children develop "islands of expertise", that is, well organized knowledge about a certain topic (e.g., birds or dinosaurs). This prior knowledge enables these children to learn new related words more easily [e.g., @chi1983].

@@

\clearpage

# Appendix A: Analyses using different phnonological distances

In the methods section, we based the choice of setting the threshold of edit distance at 2 on the fact that the early lexicon is very sparse in terms of phonological neighborhood; the early proposal that set the threshold at 1 (e.g., Vitevitch, 2008) was defined in the context of rather mature, dense lexicon. Increasing the threshold from 1 to 2 allows for a more reasonable representation of the similarity space of the early phonological network.

That said, it is useful to include the results obtained with both thresholds. In addition, it could be useful to compare the results to the case of weighted networks, i.e., with no thresholding. The main analyses for these two cases are shown in what follows. In summary, both methods of constructing the phonological networks had issues. While the fact of using edit = 1 created sparsity, using weighted networks created colinearity with length.

## Analyses using phonological networks constructed with an edit distance of 1

We show in Figure \@ref(fig:corrPlot) the correlation between the phonological connectivity and age of acquisition in both comprehension and production. The sparsity issue, due the low phonological neighborhood in the children's lexicon is apparent: most words had 0 degree and a few had non-zero but small degrees. The values of the correlations are much lower than the ones obtained with the threshold of 2.

The next figures show how the phonology-based mechanism of growth (phonoEXT) fares in comaprision to semEXT and other predictors of learing in each language (Figure \@ref(fig:staticPredEdit1) and across all languages (Figure \@ref(fig:staticAllEdit)). These figures show that phonoEXT based on edit distance 1 had no noticeable effect on learning.

```{r}
#Import data from saved
data_all <- feather::read_feather("../saved_data/data_static_scaled_edit1.feather")

data_all$language <- plyr::mapvalues(data_all$language, 
                                 from = c("Croatian","Danish","English (American)", "French (Quebecois)", "Italian", "Norwegian", "Russian", "Spanish (Mexican)", "Swedish", "Turkish"), 
                                 to = c("Croatian","Danish","English", "French", "Italian", "Norwegian", "Russian", "Spanish", "Swedish", "Turkish"))

data_long <- data_all %>%
  rename(Sem = sem_deg, Phono=phono_deg) %>%
  gather(predictor, value, Sem:frequency) %>%
  dplyr::filter(predictor == "Phono")

plot_correlation_1 <- ggplot(subset(data_long, data_long$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), 
                                  aes(x=value, y=aoa))+
  facet_grid(language~ measure)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations, correlations$language %in% c('Croatian', 'Danish', 'English', 'French','Italian')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

plot_correlation_2 <- ggplot(subset(data_long,  data_long$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(x=value, y=aoa))+
  facet_grid(language~ measure)+#, scales = "free") +
  geom_jitter(#aes(colour = lexical_class),
    colour = solarized_palette(1),
    size = 0.5, alpha = 0.5) +
  geom_abline(slope = -1)+
  coord_cartesian(xlim=c(-1,5))+
  #scale_x_continuous(limits=c(-2,5))+
  scale_y_continuous(breaks =c(15,25,35))+
  geom_smooth(method = "lm", colour = "grey3", se=FALSE)+
  scale_colour_solarized(name = "") +
  theme_few()+
  #theme_bw()+
  theme(aspect.ratio = 0.7, 
        plot.margin=grid::unit(c(0,0,0,0), "mm")
        )+
  geom_text(data=subset(correlations,  correlations$language %in% c('Norwegian', 'Russian', 'Spanish', 'Swedish','Turkish')), aes(label=paste("r=", cor, sep="")), x=3.5, y=33, size=2, fontface = "bold")+
  xlab("degree z-score") +ylab("Age (in months)")

corr_plot <- cowplot::plot_grid(
  plot_correlation_1 ,
  plot_correlation_2 ,
  nrow = 1#, 
  #align = 'vh',
  #hjust = -1
)

cor_phon_len_edit1 <- data_all %>%
  group_by(language, measure) %>%
  summarise(phon_len = cor(phono_deg, length)) %>%
  group_by(measure) %>%
  summarise(ave = mean(phon_len))
